To Batch or Not to Batch: Sequential vs. Batched Testing Strategies in the ED

### 2023-11-11 manuscript/manuscript.Rmd
adding additional results related to IV

In this paper, we use data from over 41,000 patient visits to the ED that occur during our study period to quantify the benefits and consequences of batching versus sequentially ordering diagnostic tests on patient length of stay, re-admission, and resource utilization. Our empirical strategy exploits random assignment of patients to ED physicians who differ in their propensity to batch-order diagnostic tests. When patients arrive at the ED, they are assigned to a physician based on availability, with no discretion on either side. Thus, patients who arrive at the ED at similar times are randomly assigned to physicians who vary in their willingness to batch order diagnostic tests. We measure physician tendency to batch using a leave-out, residualized measure based on all other patients the physician has seen in the ED in the study period. The tendency measure strongly predicts the ED test batch outcome but is uncorrelated with patient and ED visit characteristics.

@@

Our research design focuses on adults who visit the Mayo Clinic of Arizona ED. We observe approximately 45,000 such visits until during the stufy period. To improve power, we drop encounters with rare chief complaints (< 1000 total encounters of this kind) and complaints where a batch order occurs less than 10 percent of the time. Since batch orders are rare for these cases, our physician batch tendency instrument could suffer from a weak instrument problem were we to include them. Example complaints dropped include urinary complaints and _____. Excluding these conditions does not introduce selection bias only if physician test batching tendency is orthogonal to physician diagnosing behavior. While this assumption may be violated if we were to use a very detailed level of chief complaint information upon which to base our exclusion criterion, it is plausibly satisfied when using broad complaint categories\footnote{The idea being that, for a patient who presents with an ankle fracture, different physicians may choose different detailed diagnosis codes within the three-digit ICD-9 824 (“Fracture of ankle”), but are unlikely to disagree at the broader chief complaint cateogry.}.

@@

In our instrumental variables (IV) analysis, the key explanatory variable, $Tendency_i$, is an indicator for whether patient $i$ has their tests batch ordered at their ED encounter. This variable is crucial for understanding the causal effects of batch testing on various health outcomes and resource utilization. Below, we detail our main outcomes: (a) ED length of stay (LOS), (b) resource utilization, and (c) 72-hour return.

@@

\begin{table}[ht]
\centering
\caption{First Stage Regression Results}
\label{tab:regression}
\begin{tabular}{p{9cm}ccc}
\toprule
\textbf{} & \textbf{Model 1} & \textbf{Model 2} & \textbf{Model 3} \\
\midrule
Batching Tendency & 1.05 *** & 1.04 *** & 1.04 *** \\
 & (0.01) & (0.01) & (0.01) \\
\midrule
F statistic (full model) & 7.02 & 50.34 & 49.31  \\  
F (full model): p-value & $<0.001$ & $<0.001$ & $<0.001$ \\
\midrule
Num. obs. & 43327 & 43327 & 43327 \\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & No & Yes & Yes \\
Patient observables? & No & No & Yes \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the first stage for the baseline sample described in the text. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Column 3 corresponds to the baseline controls. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

@@

\begin{table}[htbp]
\centering
\caption{Reduced Form Regression Results with Scaled Coefficients}
\label{tab:reducedform}
\begin{tabular}{p{6cm}ccc}
\toprule
& \multicolumn{1}{c}{\textbf{Log ED LOS}}
& \multicolumn{1}{c}{\textbf{Number of Tests}}
& \multicolumn{1}{c}{\textbf{72hr Return}} \\
& \multicolumn{1}{c}{(1)}
& \multicolumn{1}{c}{(2)}
& \multicolumn{1}{c}{(3)} \\
\midrule
\textit{Model 1} & & & \\
Batch Tendency      & $0.138^{***}$ & $0.270^{***}$ & $-0.004$  \\
                    & $(0.047)$  & $(0.052)$     & $(0.003)$ \\
\\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
Hospital volume? & Yes & Yes & Yes \\
\midrule
\midrule
\textit{Model 2} & & & \\
Batch Tendency      & $-0.018$      & $0.019^{**}$ & $0.003$   \\
                    & $(0.071)$     & $(0.007)$     & $(0.003)$ \\
                    &     &     &  \\
Testing Inclination    & $0.595^{**}$  & $0.962^{***}$ & $-0.027^{**}$ \\
                    & $(0.234)$  & $(0.021)$     & $(0.011)$ \\
\\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
Hospital volume? & Yes & Yes & Yes \\
\midrule
Mean number of encounters per physician & 1805.3 &  &  \\
\midrule
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the reduced form are for the baseline sample described in the text. Coefficients on batch tendency and testing inclination are scaled by the difference between batch tendency and testing inclination respectively of the ninetieth and tenth percentile physicians for interpretability. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

@@

\begin{table}[!htbp] \centering 
  \caption{IV Results} 
  \label{} 
\begin{tabular}{p{6cm}ccc}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & ln\_ED\_LOS & RTN\_72\_HR & nEDTests \\ 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
 Batch & $-$0.097 & 0.019 & 0.113$^{***}$ \\ 
  & (0.435) & (0.020) & (0.043) \\ 
  & & & \\ 
 Testing Inclination & 0.607$^{**}$ & $-$0.027$^{**}$ & 0.964$^{***}$ \\ 
  & (0.238) & (0.011) & (0.020) \\ 
  & & & \\ 
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
Hospital volume? & Yes & Yes & Yes \\
\midrule
\textit{N} & 43,328 & 43,328 & 43,328 \\ 
R$^{2}$ & 0.195 & 0.009 & 0.306 \\ 
Adjusted R$^{2}$ & 0.190 & 0.002 & 0.302 \\ 
Residual Std. Error (df = 43061) & 0.498 & 0.187 & 0.818 \\ 
\\
\midrule
\midrule
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the IV model are for the baseline sample described in the text. Coefficients on testing inclination are scaled by the difference between testing inclination of the ninetieth and tenth percentile physicians for interpretability. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

@@

@@

>>>>

In this paper, we use data from over 45,000 patient visits to the ED that occur during our study period to quantify the benefits and consequences of batching versus sequentially ordering diagnostic tests on patient length of stay, re-admission, and resource utilization. Our empirical strategy exploits random assignment of patients to ED physicians who differ in their propensity to batch-order diagnostic tests. When patients arrive at the ED, they are assigned to a physician based on availability, with no discretion on either side. Thus, patients who arrive at the ED at similar times are randomly assigned to physicians who vary in their willingness to batch order diagnostic tests. We measure physician tendency to batch using a leave-out, residualized measure based on all other patients the physician has seen in the ED in the study period. The tendency measure strongly predicts the ED test batch outcome but is uncorrelated with patient and ED visit characteristics.

@@

Our research design focuses on adults who visit the Mayo Clinic of Arizona ED. We observe approximately 45,000 such visits until during the stufy period. To improve power, we drop encounters with rare chief complaints (< 1000 total encounters of this kind) and complaints where a batch order occurs less than 10 percent of the time. Since batch orders are rare for these cases, our physician batch tendency instrument could suffer from a weak instrument problem were we to include them. Example complaints dropped include urinary complaints and _____. Excluding these conditions does not introduce selection bias only if physician test batching tendency is orthogonal to physician diagnosing behavior. While this assumption may be violated if we were to use a very detailed level of chief complaint information upon which to base our exclusion criterion, it is plausibly satisfied when using broad complaint categories\footnote{The idea being that, for a patient who presents with an ankle fracture, different physicians may choose different detailed diagnosis codes within the three-digit ICD-9 824 (``Fracture of ankle"), but are unlikely to disagree at the broader chief complaint cateogry.}.

@@

In our instrumental variables (IV) analysis, the key explanatory variable, $Batched_i$, is an indicator for whether patient $i$ has their tests batch ordered at their ED encounter. This variable is crucial for understanding the causal effects of batch testing on various patient outcomes and resource utilization. Below, we detail our main outcomes: (a) ED length of stay (LOS), (b) resource utilization, and (c) 72-hour return.

@@

\begin{table}[ht]
\centering
\caption{First Stage: Effect of Batch Tendency on Test Batching}
\label{tab:regression}
\begin{tabular}{p{9cm}ccc}
\toprule
\textbf{} & \textbf{Model 1} & \textbf{Model 2} & \textbf{Model 3} \\
\midrule
Batching Tendency & 1.05 *** & 1.04 *** & 1.04 *** \\
 & (0.01) & (0.01) & (0.01) \\
\midrule
F statistic (full model) & 7.02 & 50.34 & 49.31  \\  
F (full model): p-value & $<0.001$ & $<0.001$ & $<0.001$ \\
\midrule
Num. obs. & 43327 & 43327 & 43327 \\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & No & Yes & Yes \\
Patient observables? & No & No & Yes \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the first stage for the baseline sample described in the text. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Column 3 corresponds to the baseline controls. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

@@

\begin{table}[htbp]
\centering
\caption{Reduced-Form Results: Length of Stay, Number of Tests, and 72-Hour Return}
\label{tab:reducedform}
\begin{tabular}{p{6cm}ccc}
\toprule
& \multicolumn{1}{c}{\textbf{Log LOS}}
& \multicolumn{1}{c}{\textbf{Number of Tests}}
& \multicolumn{1}{c}{\textbf{72-Hour Return}} \\
& \multicolumn{1}{c}{(1)}
& \multicolumn{1}{c}{(2)}
& \multicolumn{1}{c}{(3)} \\
\midrule
\textit{Model 1} & & & \\
Batch Tendency      & $0.138^{***}$ & $0.270^{***}$ & $-0.004$  \\
                    & $(0.047)$  & $(0.052)$     & $(0.003)$ \\
\\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
Hospital volume? & Yes & Yes & Yes \\
\midrule
\midrule
\textit{Model 2} & & & \\
Batch Tendency      & $-0.018$      & $0.019^{**}$ & $0.003$   \\
                    & $(0.071)$     & $(0.007)$     & $(0.003)$ \\
                    &     &     &  \\
Testing Inclination    & $0.595^{**}$  & $0.962^{***}$ & $-0.027^{**}$ \\
                    & $(0.234)$  & $(0.021)$     & $(0.011)$ \\
\\
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
Hospital volume? & Yes & Yes & Yes \\
\midrule
Mean number of encounters per physician & 1805.3 &  &  \\
\midrule
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the reduced form are for the baseline sample described in the text. Coefficients on batch tendency and testing inclination are scaled by the difference between batch tendency and testing inclination respectively of the ninetieth and tenth percentile physicians for interpretability. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

@@

In this section we examine the causal effects of batch ordering in the ED. Mirroring our presentation of the reduced-form results, we turn to presenting effects in Table \ref{tab:2SLS} on our main outcomes of ED LOS, number of tests, and 72-hour return.

\begin{table}[!htbp] \centering 
  \caption{2SLS Results: Length of Stay, Number of Tests, and 72-Hour Return} 
  \label{tab:2SLS} 
\begin{tabular}{p{6cm}ccc}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & Log LOS & Number of Tests & 72 Hour Return \\ 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
 Batch & $-$0.086 & 0.135$^{***}$ & 0.018 \\ 
  & (0.419) & (0.046) & (0.018) \\ 
  & & & \\ 
 Testing Inclination & 0.210$^{**}$ & 0.335$^{***}$ & $-$0.009$^{***}$ \\ 
  & (0.082) & (0.008) & (0.003) \\ 
  & & & \\ 
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
Hospital volume? & Yes & Yes & Yes \\
\midrule
\textit{N} & 41,929 & 41,929 & 41,929 \\ 
R$^{2}$ & 0.199 & 0.312 & 0.008 \\ 
Adjusted R$^{2}$ & 0.194 & 0.307 & 0.002 \\ 
Residual Std. Error (df = 41667) & 0.496 & 0.817 & 0.184 \\ 
\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the IV model are for the baseline sample described in the text. Coefficients on testing inclination are scaled by the difference between testing inclination of the ninetieth and tenth percentile physicians for interpretability. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

Length of Stay.—We find that batch ordering has no effect on reducing length of stay. This estimate is robust to inclusion and exclusion of controls for baseline characteristics. This implies that the anticipated efficiency gains from ordering tests in a batch, as opposed to waiting for the results of one test to come back before ordering a subsequent test, are likely washed out by the inefficiencies associated with ordering additional unnecessary tests.

Number of Tests.—Our 2SLS findings on the resource utilization are presented in column 2 of Table \ref{tab:2SLS}. We find large effects. The effect of tests batching was associated

\begin{table}[!htbp] \centering 
  \caption{2SLS Results: Specific Tests} 
  \label{tab:2SLStests} 
\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & X-Ray & Ultrasound & Non-Con CT & Con CT & Lab \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
\hline \\[-1.8ex] 
 Batch & 0.288$^{**}$ & $-$0.003 & 0.095 & $-$0.242$^{***}$ & $-$0.003 \\ 
  & (0.126) & (0.051) & (0.087) & (0.056)  & (0.080) \\ 
  & & & & &\\ 
 Testing Inclination & 0.240$^{***}$ & 0.086$^{***}$ & 0.096$^{**}$ & 0.253$^{***}$ & 0.286$^{***}$ \\ 
  & (0.056) & (0.024) & (0.039) & (0.027) & (0.038) \\ 
  & & & & & \\ 
Test Ordering Frequency & 0.47 & 0.11 & 0.21 & 0.19 & 0.77 \\ 
  & & & & &\\ 
\textit{N} & 41,929 & 41,929 & 41,929 & 41,929 & 41,929 \\ 
R$^{2}$ & 0.350 & 0.077 & 0.236 & 0.017 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the IV model are for the baseline sample described in the text. Coefficients on testing inclination are scaled by the difference between testing inclination of the ninetieth and tenth percentile physicians for interpretability. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

@@

The 2SLS results presented in the previous sections are robust to several alter- native specifications probing the key identifying assumptions of independence, exclusion, and monotonicity. We describe each robustness check in detail in online Appendix D and provide a brief overview here. To probe the independence assump- tion, we perform two checks. First, we take into account potential endogeneity in diagnosing behavior by no longer excluding any diagnosis from our analysis sample and by replacing the original diagnosis control with the most recent diag- nosis code observed prior to the ED visit (results are presented in online Appendix Table G.8, column 2). Second, we replace the physician leniency instrument with a shift-level instrument that captures the average leniency across all physicians on shift (column 3). To provide robustness checks to potential exclusion restriction violations, we adjust for chief observable margins of care that may be correlated with leniency, namely (i) decision to admit patients to the hospital, (ii) intensity of procedures performed, (iii) quality of care provided, and (iv) amount of opi- oids prescribed—i.e., intensive margin (columns 4–6). To address monotonicity, we construct physician-diagnosis-specific leniency instruments (column 7), and we check whether our instrument has a positive first stage across key demographic- and health-related subsamples (online Appendix Table G.12).

Across all robustness checks and main outcomes, we find magnitudes of our estimates broadly unchanged. Precision of estimates remains unchanged for the outcomes of long-term use and opioid-seeking behavior, while estimates for the outcomes of opioid use disorder and opioid overdose mortality become noisier in some specifications.

@@

\newpage

\begin{table}[htbp]
\centering
\caption{First Stage over all Subsamples}
\label{tab:monotonicitycheck}
\begin{tabular}{p{7cm}cccc}
\toprule
Complaint & \multicolumn{2}{c}{Batch Tendency} & N & R$^2$ \\
\midrule
Extremity Complaints & 0.662$^{***}$ & (0.079) & 5,259 & 0.126 \\
Upper Respiratory Symptoms & 1.205$^{***}$ & (0.445) & 1,915 & 0.262 \\
Abnormal Test Results & 0.670$^{**}$ & (0.325) & 1,816 & 0.141 \\
Falls, Motor Vehicle Crashes, and Trauma & 0.640$^{**}$ & (0.264) & 2,381 & 0.220 \\
Shortness of Breath & 1.574$^{***}$ & (0.468) &  2,963 & 0.147 \\
Gastrointestinal Issues & 1.080$^{***}$ & (0.241) & 3,323 & 0.094 \\
Chest Pain & 0.790$^{***}$ & (0.253) & 3,521 & 0.140 \\
Dizziness/Lightheadedness/Syncope & 1.343$^{***}$ & (0.220) & 1,894 & 0.208 \\
Back or Flank Pain & 0.859$^{**}$ & (0.385) &  2,550 & 0.123 \\
Neurological Issue & 0.416$^{*}$ & (0.232) &  3,492 & 0.232 \\
Fatigue and Weakness & 1.728$^{***}$ & (0.414) & 1,494 & 0.193 \\
Cardiac Arrhythmias & 2.432$^{***}$ & (0.548) & 1,044 & 0.327 \\
Abdominal Complaints & 1.775$^{***}$ & (0.350) & 6,231 & 0.086 \\
Skin Complaints & 0.205 & (0.212) & 2,176 & 0.133 \\
Fevers, Sweats or Chills & 1.650$^{***}$ & (0.576) &  1,842 & 0.211 \\
\bottomrule
\end{tabular}
\end{table}

### 2023-11-11 manuscript/manuscript.tex
adding additional results related to IV

Our research design focuses on adults who visit the Mayo Clinic of
Arizona ED. We observe approximately 45,000 such visits until during the
stufy period. To improve power, we drop encounters with rare chief
complaints (\textless{} 1000 total encounters of this kind) and
complaints where a batch order occurs less than 10 percent of the time.
Since batch orders are rare for these cases, our physician batch
tendency instrument could suffer from a weak instrument problem were we
to include them. Example complaints dropped include urinary complaints
and \_\_\_\_\_. Excluding these conditions does not introduce selection
bias only if physician test batching tendency is orthogonal to physician
diagnosing behavior. While this assumption may be violated if we were to
use a very detailed level of chief complaint information upon which to
base our exclusion criterion, it is plausibly satisfied when using broad
complaint
categories\footnote{The idea being that, for a patient who presents with an ankle fracture, different physicians may choose different detailed diagnosis codes within the three-digit ICD-9 824 (“Fracture of ankle”), but are unlikely to disagree at the broader chief complaint cateogry.}.

@@

In our instrumental variables (IV) analysis, the key explanatory
variable, \(Tendency_i\), is an indicator for whether patient \(i\) has
their tests batch ordered at their ED encounter. This variable is
crucial for understanding the causal effects of batch testing on various
health outcomes and resource utilization. Below, we detail our main
outcomes: (a) ED length of stay (LOS), (b) resource utilization, and (c)
72-hour return.

@@

\begin{table}[!htbp] \centering 
  \caption{IV Results} 
  \label{} 
\begin{tabular}{p{6cm}ccc}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & ln\_ED\_LOS & RTN\_72\_HR & nEDTests \\ 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
 Batch & $-$0.097 & 0.019 & 0.113$^{***}$ \\ 
  & (0.435) & (0.020) & (0.043) \\ 
  & & & \\ 
 Testing Inclination & 0.607$^{**}$ & $-$0.027$^{**}$ & 0.964$^{***}$ \\ 
  & (0.238) & (0.011) & (0.020) \\ 
  & & & \\ 
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
Hospital volume? & Yes & Yes & Yes \\
\midrule
\textit{N} & 43,328 & 43,328 & 43,328 \\ 
R$^{2}$ & 0.195 & 0.009 & 0.306 \\ 
Adjusted R$^{2}$ & 0.190 & 0.002 & 0.302 \\ 
Residual Std. Error (df = 43061) & 0.498 & 0.187 & 0.818 \\ 
\\
\midrule
\midrule
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the IV model are for the baseline sample described in the text. Coefficients on testing inclination are scaled by the difference between testing inclination of the ninetieth and tenth percentile physicians for interpretability. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

@@

>>>>

Our research design focuses on adults who visit the Mayo Clinic of
Arizona ED. We observe approximately 45,000 such visits until during the
stufy period. To improve power, we drop encounters with rare chief
complaints (\textless{} 1000 total encounters of this kind) and
complaints where a batch order occurs less than 10 percent of the time.
Since batch orders are rare for these cases, our physician batch
tendency instrument could suffer from a weak instrument problem were we
to include them. Example complaints dropped include urinary complaints
and \_\_\_\_\_. Excluding these conditions does not introduce selection
bias only if physician test batching tendency is orthogonal to physician
diagnosing behavior. While this assumption may be violated if we were to
use a very detailed level of chief complaint information upon which to
base our exclusion criterion, it is plausibly satisfied when using broad
complaint
categories\footnote{The idea being that, for a patient who presents with an ankle fracture, different physicians may choose different detailed diagnosis codes within the three-digit ICD-9 824 (``Fracture of ankle"), but are unlikely to disagree at the broader chief complaint cateogry.}.

@@

In our instrumental variables (IV) analysis, the key explanatory
variable, \(Batched_i\), is an indicator for whether patient \(i\) has
their tests batch ordered at their ED encounter. This variable is
crucial for understanding the causal effects of batch testing on various
patient outcomes and resource utilization. Below, we detail our main
outcomes: (a) ED length of stay (LOS), (b) resource utilization, and (c)
72-hour return.

@@

In this section we examine the causal effects of batch ordering in the
ED. Mirroring our presentation of the reduced-form results, we turn to
presenting effects in Table \ref{tab:2SLS} on our main outcomes of ED
LOS, number of tests, and 72-hour return.

\begin{table}[!htbp] \centering 
  \caption{2SLS Results: Length of Stay, Number of Tests, and 72-Hour Return} 
  \label{tab:2SLS} 
\begin{tabular}{p{6cm}ccc}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & Log LOS & Number of Tests & 72 Hour Return \\ 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
 Batch & $-$0.086 & 0.135$^{***}$ & 0.018 \\ 
  & (0.419) & (0.046) & (0.018) \\ 
  & & & \\ 
 Testing Inclination & 0.210$^{**}$ & 0.335$^{***}$ & $-$0.009$^{***}$ \\ 
  & (0.082) & (0.008) & (0.003) \\ 
  & & & \\ 
Seasonality and shift fixed effects? & Yes & Yes & Yes \\
Chief Complaint? & Yes & Yes & Yes \\
Patient observables? & Yes & Yes & Yes \\
Hospital volume? & Yes & Yes & Yes \\
\midrule
\textit{N} & 41,929 & 41,929 & 41,929 \\ 
R$^{2}$ & 0.199 & 0.312 & 0.008 \\ 
Adjusted R$^{2}$ & 0.194 & 0.307 & 0.002 \\ 
Residual Std. Error (df = 41667) & 0.496 & 0.817 & 0.184 \\ 
\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the IV model are for the baseline sample described in the text. Coefficients on testing inclination are scaled by the difference between testing inclination of the ninetieth and tenth percentile physicians for interpretability. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

Length of Stay.---We find that batch ordering has no effect on reducing
length of stay. This estimate is robust to inclusion and exclusion of
controls for baseline characteristics. This implies that the anticipated
efficiency gains from ordering tests in a batch, as opposed to waiting
for the results of one test to come back before ordering a subsequent
test, are likely washed out by the inefficiencies associated with
ordering additional unnecessary tests.

Number of Tests.---Our 2SLS findings on the resource utilization are
presented in column 2 of Table \ref{tab:2SLS}. We find large effects.
The effect of tests batching

\begin{table}[!htbp] \centering 
  \caption{2SLS Results: Specific Tests} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & X-Ray & Ultrasound & Non-Con CT & Con CT & Lab \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
\hline \\[-1.8ex] 
 Testing Inclination & 0.240$^{***}$ & 0.086$^{***}$ & 0.096$^{**}$ & 0.253$^{***}$ & 0.286$^{***}$ \\ 
  & (0.056) & (0.024) & (0.039) & (0.027) & (0.038) \\ 
  & & & & & \\ 
 Batch & 0.288$^{**}$ & $-$0.003 & 0.095 & $-$0.242$^{***}$ & $-$0.003 \\ 
  & (0.126) & (0.051) & (0.087) & (0.056)  & (0.080) \\ 
  & & & & &\\ 
Test Ordering Frequency & 0.47 & 0.11 & 0.21 & 0.19 & 0.77 \\ 
  & & & & &\\ 
\textit{N} & 41,929 & 41,929 & 41,929 & 41,929 & 41,929 \\ 
R$^{2}$ & 0.350 & 0.077 & 0.236 & 0.017 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Estimates of the IV model are for the baseline sample described in the text. Coefficients on testing inclination are scaled by the difference between testing inclination of the ninetieth and tenth percentile physicians for interpretability. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Patient observables include sex dummy, race/ethnicity, and age bins. Robust standard errors are clustered at the physician level.}
\item $*** p < 0.001$, $** p < 0.01$, $* p < 0.05$.
\end{tablenotes}
\end{table}

@@

The 2SLS results presented in the previous sections are robust to
several alter- native specifications probing the key identifying
assumptions of independence, exclusion, and monotonicity. We describe
each robustness check in detail in online Appendix D and provide a brief
overview here. To probe the independence assump- tion, we perform two
checks. First, we take into account potential endogeneity in diagnosing
behavior by no longer excluding any diagnosis from our analysis sample
and by replacing the original diagnosis control with the most recent
diag- nosis code observed prior to the ED visit (results are presented
in online Appendix Table G.8, column 2). Second, we replace the
physician leniency instrument with a shift-level instrument that
captures the average leniency across all physicians on shift (column 3).
To provide robustness checks to potential exclusion restriction
violations, we adjust for chief observable margins of care that may be
correlated with leniency, namely (i) decision to admit patients to the
hospital, (ii) intensity of procedures performed, (iii) quality of care
provided, and (iv) amount of opi- oids prescribed---i.e., intensive
margin (columns 4--6). To address monotonicity, we construct
physician-diagnosis-specific leniency instruments (column 7), and we
check whether our instrument has a positive first stage across key
demographic- and health-related subsamples (online Appendix Table G.12).

Across all robustness checks and main outcomes, we find magnitudes of
our estimates broadly unchanged. Precision of estimates remains
unchanged for the outcomes of long-term use and opioid-seeking behavior,
while estimates for the outcomes of opioid use disorder and opioid
overdose mortality become noisier in some specifications.
